---
title: "CCGeno"
author: "LucianoRogerio; Caroline Cardoso"
date: "2021-10-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Coleções Nucleares Genotípicas

```{r Lendo arquivo de Dados Genotípicos e Fenotípicos, eval = F}
library(here)
library(corehunter); library(tidyverse); library(data.table)
DataCar <- readRDS(file = here::here("output", "DataSelPreparedCar.rds"))

GBSDataCar2 <- readRDS(file = here::here("data", "DadosGBSCaroline2.rds"))

CHCar <- genotypes(GBSDataCar2, format = "biparental")
CCSize <- 331
```


### Coleção Nuclear 1 - Shannon Weaver optimization with Modified Rogers Distance

```{r SH/MR, eval = F}
CCC4 <- sampleCore(CHCar, objective(type = "SH", measure = "MR"), size = CCSize)
CloneSel <- CCC4$sel

DataCarCCGeno4 <- DataCar[DataCar$Acessos%in%CloneSel,]

write.table(DataCarCCGeno4,
            file= here::here("output", "CCGenData", "DadosGeno1CCGeno1.CSV"),
            quote = F, sep = ";", dec = ".")
```

### Coleção Nuclear 2 - Shannon optimization with Cavalli-Sforza and Edwards Distance

```{r SH/CE, eval = F}
CCC5 <- sampleCore(CHCar, objective(type = "SH", measure = "CE"), size = CCSize)

DataCarCCGeno5 <- DataCar[DataCar$Acessos%in%CloneSel,]

write.table(DataCarCCGeno5,
            file = here::here("output", "CCGenData", "DadosGeno2CCGeno1.CSV"),
            quote = F, sep = ";", dec = ".")
```

### Coleção Nuclear 3 - Expected Proportion of Heterozigous Loci optimization with Modified Rogers Distance

```{r HE/MR, eval = F}
CCC6 <- sampleCore(CHCar, objective(type = "HE", measure = "MR"), size = CCSize)
CloneSel <- CCC6$sel 

DataCarCCGeno6 <- DataCar[DataCar$Acessos%in%CloneSel,]

write.table(DataCarCCGeno6,
            file = here::here("output", "CCGenData", "DadosGeno3CCGeno1.CSV"), quote = F, sep = ";", dec = ".")
```

### Coleção Nuclear 4 - Expected Proportion of Heterozigous Loci optimization with Cavalli-Sforza and Edwards Distance

```{r HE/CE, eval = F}
CCC7 <- sampleCore(CHCar, objective(type = "HE", measure = "CE"), size = CCSize)
CloneSel <- CCC7$sel

DataCarCCGeno7 <- DataCar[DataCar$Acessos%in%CloneSel,]

write.table(DataCarCCGeno7,
            file= here::here("output", "CCGenData", "DadosGeno4CCGeno1.CSV"), quote = F, sep = ";", dec = ".")
```

### Coleção Nuclear 5 - Allele coverage optimization with Modified Rogers Distance

```{r CV/MR, eval = F}
CCC8 <- sampleCore(CHCar, objective(type = "CV", measure = "MR"), size = CCSize)
CloneSel <- CCC8$sel

DataCarCCGeno8 <- DataCar[DataCar$Acessos%in%CloneSel,]

write.table(DataCarCCGeno8,
            file= here::here("output", "CCGenData", "DadosGeno5CCGeno1.CSV"), quote = F, sep = ";", dec = ".")
```

### Coleção Nuclear 6 - Allele coverage optimization with Cavalli-Sforza and Edwards Distance

```{r CV/CE, eval = F}
CCC9 <- sampleCore(CHCar, objective(type = "CV", measure = "CE"), size = CCSize)
CloneSel <- CCC9$sel

DataCarCCGeno9 <- DataCar[DataCar$Acessos%in%CloneSel,]

write.table(DataCarCCGeno9, file = here::here("output", "CCGenData", "DadosGeno6CCGeno1.CSV"), quote = F, sep = ";", dec = ".")
```

### Coleção Nuclear 7 - Allele Match Distance for MLST - DARwin software

```{r AM/MLST DARwin, eval = F}
CCSelAM.DARwin <- read.table(here::here("output", "CarDarwinMatrix.don"),
                             sep = "\t", header = T,skip = 2,
                             nrows = 1363)

ClonesSel <- CCSelAM.DARwin %>% filter(CCSel == "Kept") %>%
  select(N)
DataCarCCGeno10 <- DataCar[DataCar$Acessos%in%ClonesSel$N,]

write.table(DataCarCCGeno10, file= here::here("output", "CCGenData", "DadosGeno7CCGeno1.CSV"), quote = F, sep = ";", dec = ".")
```

### Coleção Nuclear 8 - Modified Rodgers distances for MLST - DARwin software

```{r MR/MLST DARwin, eval = F}
library(poppr); library(adegenet); library(here); library(tidyverse)
GenIdCar <- df2genind(GBSDataCar2, ploidy = 2, type = "PA", ncode = 1)


MRDist <- rogers.dist(GenIdCar)

cat(nrow(as.matrix(MRDist)), "\n", file = here::here("output", "MatrizMDCaroline.phy"))
write.table(x = as.matrix(MRDist), file = here::here("output", "MatrizMDCaroline.phy"),
          append = TRUE, quote = FALSE, sep = "\t", col.names = FALSE)

```

Rodar após concluir a análise de MLST no DARwin
```{r MR/MLST results, eval = F}
CCSelMR.DARwin <- read.table(here::here("output", "MatrizMDCaroline.DON"),
                             sep = "\t", header = T,skip = 2,
                             nrows = 1336)
ClonesSel <- CCSelMR.DARwin %>% filter(CCSel == "Kept") %>%
  select(Name)
DataCarCCGeno11 <- DataCar[DataCar$Acessos%in%ClonesSel$Name,]

write.table(DataCarCCGeno11, file= here::here("output", "CCGenData", "DadosGeno8CCGeno1.CSV"), quote = F, sep = ";", dec = ".")

```

### Coleção Nuclear 9 - Cavalli-Sforza and Edwards Distance for MLST - DARwin software

```{r CE/MLST DARwin, eval = F}
library(poppr); library(adegenet); library(here); library(phangorn)
library(data.table)
CEDist <- edwards.dist(GenIdCar)

cat(nrow(as.matrix(CEDist)), "\n", file = here::here("output", "MatrizCECaroline.phy"))
write.table(x = as.matrix(CEDist), file = here::here("output", "MatrizCECaroline.phy"),
          append = TRUE, quote = FALSE, sep = "\t", col.names = FALSE)
```

Rodar após concluir a análise de MLST no DARwin
```{r CE/MLST results, eval = F}
CCSelCE.DARwin <- read.table(here::here("output","MatrizCECaroline.DON"),
                             sep = "\t", header = T, skip = 2,
                             nrows = 1336)
ClonesSel <- CCSelCE.DARwin %>% filter(CCSel == "Kept") %>%
  select(Name)

DataCarCCGeno12 <- DataCar[DataCar$Acessos%in%ClonesSel$Name,]

write.table(DataCarCCGeno12, file= here::here("output", "CCGenData", "DadosGeno9CCGeno1.CSV"), quote = F, sep = ";", dec = ".")

```

Back - [Coleções Nucleares Fenotípicas](CCPheno.html)

Next - [Análises Discriminantes](AnDis.html)

[Home](index.html)
