---
title: "Diversidade Genetica"
author: "LucianoRogerio"
date: "2021-11-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

```{r}
suppressMessages(library(tidyverse)); suppressMessages(library(data.table)); suppressMessages(library(reactable))
source(here::here("code", "Shannon-Weaver.R"))
Alldataset <- readRDS(here::here("output", "AllDataCCCaroline.RDS"))

QualityTrait <- colnames(Alldataset)[2:36]


AlldataSetQuant <- Alldataset %>% select(-all_of(QualityTrait))
AlldataSetQuali <- Alldataset %>% select(Acessos, QualityTrait, Data, Method) %>% suppressMessages()

QuantTrait <- data.frame(Trait = colnames(AlldataSetQuant)[2:17],
                         Min = apply(AlldataSetQuant[2:17], FUN = min, na.rm = TRUE, MARGIN = 2),
                         Max = apply(AlldataSetQuant[2:17], FUN = max, na.rm = TRUE, MARGIN = 2))
DataMethod <- unique(data.frame(Data = AlldataSetQuant$Data,
                                Method = AlldataSetQuant$Method))

QuantSHs <- NULL
for(i in 1:nrow(DataMethod)){
  for(trait in QuantTrait$Trait){
    QuantSH <- data.frame(Data = DataMethod$Data[i],
                          Method = DataMethod$Method[i],
                          Trait = trait,
                          SH = Shannon.Weaver.QT(AlldataSetQuant[(AlldataSetQuant$Data%in%DataMethod$Data[i] &
                                                                    AlldataSetQuant$Method%in%DataMethod$Method[i]),trait],
                                                 Min = QuantTrait[QuantTrait$Trait == trait, "Min"],
                                                 Max = QuantTrait[QuantTrait$Trait == trait, "Max"],
                                                 formula = 2))
    QuantSHs <- rbind(QuantSHs, QuantSH)
  }
}

write.table(QuantSHs, here::here("output", "Shannon-WeaverCCQuantIndex.csv"), dec = ".",
            sep = ",", quote = F, row.names = F)

Table1SH <- matrix(QuantSHs$SH, byrow = F, ncol = 13)
rownames(Table1SH) <- unique(QuantSHs$Trait)
colnames(Table1SH) <- unique(paste(QuantSHs$Data, QuantSHs$Method, sep = " - "))
colnames(Table1SH)[1] <- "BAG"
```


#### **Table 1.** Shannon-Weaver indexes for Cassava quantitative traits of the Bank of germplasm and 12 nuclear core collections.
```{r, echo = FALSE}
Table1SH %>% reactable(defaultPageSize = 16, columns = list(
  BAG = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Pheno - GW/CH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Pheno - CH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Pheno - GW/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/SH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/SH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/EH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/EH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/AC' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/AC' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - AM/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US"))))


Trait <- c("CorFolhaApical" = 4, "CorPeciolo" = 6, "CorFolhaDesenv" = 5,
           "CorNerv" = 4, "CorCortexCaule" = 4, "CorExternaCaule" = 7,
           "CorEpidCaule" = 4, "CorRamosTerm" = 3, "CorExternaRzs" = 4,
           "CorCortexRzs" = 4, "CorPolpaRzs" = 5, "NLobulos" = 5,
           "Pubescencia" = 2, "FormaLobuloCentral" = 9, "PosicaoPeciolo" = 4,
           "Sinuosidade" = 2, "ComprEstip" = 2, "HabCresCaule" = 2,
           "TipoPlan" = 4, "MargEstip" = 2, "HabRamif" = 4, "NiveisRam" = 4, 
           "ProemCicatrizesFolhas" = 2, "FormaRzs" = 4, "TxtEpidermeRzs" = 2,
           "PresPedunculoRzs" = 3, "PosicaoRzs" = 3, "Flowering" = 2,
           "DestqPelicRz" = 2, "DestqCortexRz" = 3, "ConstrRzs" = 3,
           "ComprFilotaxia" = 3, "AnguloRamif" = 22, "PorteMod" = 5, 
           "HCNMod" = 9)

QualTrait <- data.frame(Trait = names(Trait),
                        NClas = Trait,
                        row.names = NULL)

QualSHs <- NULL
for(i in 1:nrow(DataMethod)){
  for(trait in QualTrait$Trait) {
    QualSH <- data.frame(Data = DataMethod$Data[i],
                         Method = DataMethod$Method[i],
                         Trait = trait,
                         SH = Shannon.Weaver.QL(as.character(AlldataSetQuali[(AlldataSetQuali$Data%in%DataMethod$Data[i] &
                                                                    AlldataSetQuali$Method%in%DataMethod$Method[i]),trait]),
                                                nclass = QualTrait$NClas[QualTrait$Trait%in%trait]))
    QualSHs <- rbind(QualSHs, QualSH)
  }
}

write.table(QualSHs, here::here("output", "Shannon-WeaverCCQualIndex.csv"), dec = ".",
            sep = ",", quote = F, row.names = F)


Table2SH <- matrix(QualSHs$SH, byrow = F, ncol = 13)
rownames(Table2SH) <- unique(QualSHs$Trait)
colnames(Table2SH) <- unique(paste(QualSHs$Data, QualSHs$Method, sep = " - "))
colnames(Table2SH)[1] <- "BAG"


SHInd <- as.data.frame(rbind(Table1SH, Table2SH))
```


#### **Table 2.** Shannon-Weaver indexes for Cassava quality traits of the Bank of germplasm and 12 nuclear core collections.
```{r, echo = FALSE}
Table2SH %>% reactable(defaultPageSize = 15, columns = list(
  BAG = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Pheno - GW/CH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Pheno - CH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Pheno - GW/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/SH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/SH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/EH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/EH' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/AC' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/AC' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - AM/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - MR/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US")),
  'Geno - CE/MLST' = colDef(format = colFormat(digits = 3, locales = "en-US"))))
```


## Core Collection Selection - Phenotypic data
```{r}
SHBAG <- SHInd[, "BAG"]
SHIndPhen <- SHInd[ , colnames(SHInd)%like%"Pheno"]

SHRelPhen <- (SHIndPhen - SHBAG)*100/SHBAG

SHSelPhen <- data.frame(Method = colnames(SHRelPhen),
                        SHMeanImprov = colMeans(SHRelPhen, na.rm = TRUE),
                        SHsd = apply(SHRelPhen, FUN = sd, na.rm = TRUE, MARGIN = 2),
                        row.names = NULL)
```

#### **Table 3.** Mean improvement of the Shannon-Weaver diversity index in % of nuclear core collections estimated of Phenotypic data compared to the Bank of Germplasm
```{r, echo = FALSE}
reactable(SHSelPhen, columns = list(
  SHMeanImprov = colDef(format = colFormat(digits = 3, locales = "en-US")),
  SHsd = colDef(format = colFormat(digits = 3, locales = "en-US"))))

```

## Core Collection Selection - Genotypic data
```{r}
SHIndGen <- SHInd[ , colnames(SHInd)%like%"Geno"]

SHRelGen <- (SHIndGen - SHBAG)*100/SHBAG

SHSelGen <- data.frame(Method = colnames(SHRelGen),
                        SHMeanImprov = colMeans(SHRelGen, na.rm = TRUE),
                        SHsd = apply(SHRelGen, FUN = sd, na.rm = TRUE, MARGIN = 2),
                        row.names = NULL)
```

#### **Table 4.** Mean improvement of the Shannon-Weaver diversity index in % of nuclear core collections estimated of Genotypic data compared to the Bank of Germplasm
```{r, echo = FALSE}
reactable(SHSelGen, columns = list(
  SHMeanImprov = colDef(format = colFormat(digits = 3, locales = "en-US")),
  SHsd = colDef(format = colFormat(digits = 3, locales = "en-US"))))

```


Back - [Estimativas de Diversidade genética](DivGen.html)
[Home](index.html)
